---
title: "Lab 3 244"
author: "Danielle Hoekstra"
date: "2023-01-26"
output: html_document
---

```{r setup, echo = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(GGally)
library(jtools)
library(AICcmodavg)

# shorcut reminder command+option+i adds in an r chunk
```

# Pseudocode: a way to put the steps in even without the code

* Examine our data (plots, tables, summary tables)
* Identify a question
* Wrangle the data 
* Identify some candidate models
* Select among candidate models using AIC/BIC
* Select among candidate models using K-fold cross validation
* Select among candidate models using area under Receiver Operating Characteristic Curve

```{r}
GGally::ggpairs(penguins %>% select(species, bill_length_mm:sex),
                aes(color = species)) #colon to select all columns between two variables
```
```{r}
class(penguins$species)
levels(penguins$species)
#factors that have different levels to them

adelie_chinstrap <- penguins %>%
  filter(species %in% c('Adelie', 'Chinstrap')) %>% 
  mutate(species = fct_drop(species)) %>% 
  select(-year) %>% 
  drop_na()

#using %in% the idea is to not have the pattern be the identifier like in ==, == will set whatever is inside the c() as the repeating pattern to match (remember Casey's example on the white board) %in% tells you to keep all of what is in c() 
```
## Look at the trends across variables
```{r}
ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = flipper_length_mm)) +
  geom_point(aes(color = sex, shape = island)) +
  facet_wrap(~ species)

ggplot(data = adelie_chinstrap, aes(x = body_mass_g, y = bill_length_mm)) +
  geom_point(aes(color = sex, shape = island)) +
  facet_wrap(~ species)
```
## Binary logistic regression!

```{r}
f1 <- species ~ body_mass_g + flipper_length_mm + sex

ad_chin_blr1 <- glm(formula = f1, data = adelie_chinstrap, 
                         family = 'binomial') 
ad_chin_blr1 

summary(ad_chin_blr1)

blr1_tidy <- tidy(ad_chin_blr1) #tabular format, easier to read

#adelie is our reference value, therefore a negative value means that decreasing the body mass decreases the probability of the penguin being an Adelie
```

```{r}
ggplot(data = adelie_chinstrap, aes(x = species, y = body_mass_g)) +
  geom_jitter(aes(color = sex))
```
```{r}
blr1_fitted <- ad_chin_blr1 %>% 
  broom::augment(type.predict = 'response')
# fitted column, based on the variables, fitted says the percentage that it is a type 1, or type 2 etc...

ggplot(data = blr1_fitted, aes(x = flipper_length_mm, y = .fitted)) + 
  geom_point(aes(color = sex, shape = species)) +
  geom_smooth(aes(color = sex), se = FALSE) +
  labs(x = 'Flipper length (mm)', y = 'Probability of outcome (Chinstrap)')

#species are more similar so it may be tougher to tell a difference or to pick out a clear difference
```
## Predictions for new values with predict()
```{r}
ex1 <- predict(ad_chin_blr1, 
               data.frame(sex = "female", 
                         body_mass_g = 3410,
                         flipper_length_mm = 192),
               type = 'response')

new_df <- data.frame(
  sex = c('male', 'female', 'female'), 
  body_mass_g = c(3298, 4100, 3600),
  flipper_length_mm = c(212, 175, 180)
)

ex2 <- predict(ad_chin_blr1, new_df, type = 'response')
ex2
```
## Create a new binary logistic model
```{r}
f2 <- species ~ bill_length_mm + body_mass_g

ad_chin_blr2 <- glm(formula = f2, data = adelie_chinstrap, family = 'binomial')

ad_chin_blr2
summary(ad_chin_blr2)

blr2_tidy <- broom::tidy(ad_chin_blr2)
```

```{r}
ggplot(adelie_chinstrap, aes(x = bill_length_mm, y = body_mass_g)) +
  geom_point(aes(color = species))
```
## Model selection

```{r}
AICcmodavg::aictab(list(ad_chin_blr1, ad_chin_blr2)) 
# LL or log likelihood is driving it, the difference of 2 in the delta_AIC is good, so a difference of 200 + doesn't need to be considered

AICcmodavg::bictab(list(ad_chin_blr1, ad_chin_blr2)) 
# get to basically see the same thing, more likely to look at a cross validation approach
```

## 10-fold cross validation
```{r}
set.seed(123)

n_folds <- 10
fold_vec <- rep(1:n_folds, length.out = nrow(adelie_chinstrap))

ad_chin_kfold <- adelie_chinstrap %>%  
  mutate(fold = sample(fold_vec, size = n(), replace = FALSE))

```

## purrr::map()
```{r}
x_vec <- 1:10
thing <- purrr::map(.x = x_vec, ### a sequence(vector, list)
                    .f = sqrt) ### a function applied to every element in the sequence
thing

my_funct <- function(x, y, z) {
  return((x - y)^z)
}

thing2 <- purrr::map(.x = x_vec,
                     .f = my_funct,
                     y = 2, z = 3)
thing2
```

```{r}
pred_acc <- function(x,y) {
  accurate <- ifelse(x == y, 1, 0)
  
  return(mean(accurate, na.rm = TRUE))
}

calc_fold <- function(i, fold_df, f) {
  kfold_test <- fold_df %>% 
    filter(fold == i)
  kfold_train <- fold_df %>% 
    filter(fold != i) #basically a test group and initial group
  
  kfold_blr <- glm(f, data = kfold_train, family = 'binomial')
  
  kfold_pred <- kfold_test %>% 
    mutate(blr = predict(kfold_blr, kfold_test, type = 'response'),
           pred = ifelse(blr > 0.50, 'Chinstrap', 'Adelie'))
  
  kfold_accuracy <- kfold_pred %>% 
    summarize(blr_acc = pred_acc(species, pred)) 
  
  return(kfold_accuracy)
}

#ifelse, is if ____ then _____
#apply predicted accuracy function, two functions, actualy species and predicted species to see how accurate
```

```{r}
results1_purr_df <- purrr::map(.x = 1:n_folds,
                               .f = calc_fold,
                               fold_df = ad_chin_kfold,
                               f = f1) %>% 
  bind_rows() %>% 
  mutate(mdl = 'f1')

results2_purr_df <- purrr::map(.x = 1:n_folds,
                               .f = calc_fold,
                               fold_df = ad_chin_kfold,
                               f = f2) %>% 
  bind_rows() %>% 
  mutate(mdl = 'f2')
                       
results_purr_df <- bind_rows(results1_purr_df, results2_purr_df) %>% 
  group_by(mdl) %>% 
  summarize(mean_acc = mean(blr_acc))
results_purr_df
```
# Tidymodels version

```{r}
### define model type
blr_model <- logistic_reg() %>% 
  set_engine('glm') 

#could also change glm to brulee if that is the prefered set engine, or the company prefers one over the other

### basic regression
blr_tidyfit_f1 <- blr_model %>% 
  fit(f1, data = adelie_chinstrap)

blr_tidyfit_f2 <- blr_model %>% 
  fit(f2, data = adelie_chinstrap)

blr_tidyfit_f1 %>% 
  tidy()
blr_tidyfit_f2 %>% 
  tidy()
blr_tidyfit_f1 %>% 
  glance() #BIC and AIC
```

## More crossfold validation: Tidy kfold
```{r}
set.seed(345)

tidy_folds <- vfold_cv(adelie_chinstrap, v = 10)

### use workflow to bundle a model and a formula, built into tidymodel package, pre packaging your data, bundling it all up to you can run it easily and change bits at a time

blr_tidy_wf1 <- workflow() %>% 
  add_model(blr_model) %>% 
  add_formula(f1)

### cross validation
blr_tidy_cv_f1 <- blr_tidy_wf1 %>% 
  fit_resamples(tidy_folds) #apply this workflow to my tidy folds data frame


collect_metrics(blr_tidy_cv_f1)

#run it again with f2

blr_tidy_wf2 <- workflow() %>% 
  add_model(blr_model) %>% 
  add_formula(f2)

### cross validation
blr_tidy_cv_f2 <- blr_tidy_wf2 %>% 
  fit_resamples(tidy_folds) #apply this workflow to my tidy folds data frame


collect_metrics(blr_tidy_cv_f2) #mean is around 97%, much higher average
```

# Area under the curve
```{r}
#make predictions
blr_f1_pred <- adelie_chinstrap %>% 
  mutate(predict(blr_tidyfit_f1, .)) %>% 
  mutate(predict(blr_tidyfit_f1, ., type = 'prob'))

blr_f1_pred %>% 
  roc_curve(truth = species, .pred_Adelie) %>% 
  autoplot() 

# period means operate on the data set I have already classified
# another way to look at model prediction, sensitivity = how well does this identify Adelie penguins vs specificity = how well does it predict chinstraps to predict Adelie penguins, ideally false positives to be zero
```










